### 2019-02-13 Hopf Algebra Notes/Hopf Algebras.tex
I should push more frequently

\documentclass[12pt]{article}

\usepackage{setspace}

\usepackage{amsmath, amsfonts, amssymb, graphicx, color, fancyhdr, lipsum, scalerel, stackengine, mathrsfs, tikz-cd, mdframed, enumitem, framed, adjustbox, bm, upgreek, x	color}
\usepackage[framed,thmmarks]{ntheorem}
\usepackage[mathscr]{euscript}

%set up theorem/definition/etc envs
%Problems will be created using their own counter and style
\theoremstyle{break}
\theoreminframepreskip{0pt}
\theoreminframepostskip{0pt}
\newframedtheorem{prob}{Problem}[section]

%solution template
\theoremstyle{nonumberbreak}
\theoremindent0.5cm
\theorembodyfont{\upshape}
\theoremseparator{:}
\theoremsymbol{\ensuremath\spadesuit}
\newtheorem{sol}{Solution}

%Theorems
\definecolor{thmcol}{RGB}{0,102,34}
\theoremstyle{changebreak}
\theoremseparator{}
\theoremsymbol{}
\theoremindent0.5cm
\theoremheaderfont{\color{thmcol}\bfseries} 
\newtheorem{thm}{Theorem}[subsection]

%Lemmas and Corollaries
\theoremheaderfont{\bfseries}
\newtheorem{lem}[thm]{Lemma}
\newtheorem{cor}[thm]{Corollary}

%Create a new env that references a theorem and creates a 'primed' version
%Note this can be used recursively to get double, triple, etc primes
\newenvironment{thm-prime}[1]
  {\renewcommand{\thethm}{\ref{#1}$'$}%
   \addtocounter{thm}{-1}%
   \begin{thm}}
  {\end{thm}}

\setlength\fboxsep{15pt}

%Shade definitions
\theoremindent0cm
\theoremheaderfont{\normalfont\bfseries} 
\def\theoremframecommand{\colorbox[rgb]{.85,1,.85}}
\newshadedtheorem{defn}[thm]{Definition}

%Example
\theoremstyle{break}
\def\theoremframecommand{\colorbox[rgb]{0.9,0.9,0.9}}
\newshadedtheorem{ex}{Example}[section]

%Man, that's really good! Let's use the same thing for definitons.
\newenvironment{def-prime}[1]
  {\renewcommand{\thethm}{\ref{#1}$'$}%
   \addtocounter{thm}{-1}%
   \begin{def}}
  {\end{def}}

%proofs
\theoremstyle{nonumberbreak}
\theoremindent0.5cm
\theoremheaderfont{\sc}
\theoremseparator{}
\theoremsymbol{\ensuremath\spadesuit}
\newtheorem{prf}{Proof}

\theoremstyle{nonumberplain}
\theoremseparator{:}
\theoremsymbol{}
\newtheorem{conj}{Conjecture}

%remarks
\theoremstyle{change}
\theoremindent0.5cm
\theoremheaderfont{\sc}
\theoremseparator{:}
\theoremsymbol{}
\newtheorem{rmk}[thm]{Remark}

%Replacement for the old geometry package
\usepackage{fullpage}

%Put page breaks before each part
\let\oldpart\part%
\renewcommand{\part}{\clearpage\oldpart}%

%Center each figure by default
\makeatletter
\g@addto@macro\@floatboxreset{\centering}
\makeatother

%header stuff
\setlength{\headsep}{24pt}  % space between header and text
\pagestyle{fancy}     % set pagestyle for document
\lhead{Notes on Hopf Algebras} % put text in header (left side)
\rhead{Nico Courts} % put text in header (right side)
\cfoot{\itshape p. \thepage}
\setlength{\headheight}{15pt}
\allowdisplaybreaks

%Set of Integers
\newcommand*{\Z}{
\mathbb{Z}
}
%Set of Natural Numbers
\newcommand*{\N}{
\mathbb{N}
}
%Set of Real Numbers
\newcommand*{\R}{
\mathbb{R}
}
%Set of Complex Numbers
\newcommand*{\C}{
\mathbb{C}
}
%Rationals
\newcommand*{\Q}{
\mathbb{Q}
}

%Section break
\newcommand*{\brk}{
\rule{2in}{.1pt}
}

\DeclareMathOperator{\Aut}{Aut}

%raise that Chi!
\DeclareRobustCommand{\Chi}{{\mathpalette\irchi\relax}}
\newcommand{\irchi}[2]{\raisebox{\depth}{$#1\chi$}}

%Image
\DeclareMathOperator{\im}{Im}

\DeclareMathOperator{\Ext}{Ext}
\DeclareMathOperator{\gldim}{gldim}
\DeclareMathOperator{\projdim}{projdim}
\DeclareMathOperator{\findim}{findim}

%Coker
\DeclareMathOperator{\coker}{coker}

%characteristic
\DeclareMathOperator{\ch}{char}

%rank
\DeclareMathOperator{\rank}{rank}

%identity map
\DeclareMathOperator{\id}{id}

\newcommand*{\Hom}{\ensuremath\text{Hom}}

%Hopf algebra stuff
\newcommand*{\Vectk}{\operatorname{Vect}_k}
\newcommand*{\Algk}{\operatorname{Alg}_k}
\newcommand*{\Coalgk}{\operatorname{Coalg}_k}
\newcommand*{\calM}{\mathcal{M}}
\DeclareMathOperator{\Coalg}{Coalg}
\DeclareMathOperator{\Alg}{Alg}
\newcommand*{\scrT}{\ensuremath\mathscr{T}}
\newcommand*{\Ch}{\mathbf{Ch}}

%fix tilde
\let\tilde\relax
\newcommand*{\tilde}[1]{\widetilde{#1}}

% Enumerate will automatically use letters (e.g. part a,b,c,...)
\setenumerate[0]{label=(\alph*)}

\begin{document}
%make the title page
\title{Hopf Algebras\vspace{-1ex}}
\author{A course by: Prof. James Zhang\\
Notes by: Nico Courts}
\date{Winter 2019}
\maketitle

\renewcommand{\abstractname}{Introduction}
\begin{abstract}
	These are the notes I took in class during the Winter 2019 topics course
	\textit{Math 582H - Hopf Algebras} at University of Washington, Seattle.

	The course description follows:

	\brk

	This course is an introduction to Hopf algebras. In addition to basic material in 
	Hopf algebra, we will present some latest developments in quantum groups and tensor 
	and fusion categories. One of the newer topics is homological properties of 
	Noetherian Hopf algebras of low Gelfand-Kirillov dimension. A good reference for 
	the first two topics in the book \textit{Hopf Algebras and Their Action Rings} by 
	Susan Montgomery. Here is a list of possible topics:

\begin{itemize}
	\item Classical theorems concerning finite dimensional Hopf algebras.
	\item Infinite dimensional Hopf algebras and quantum groups.
	\item Duality and Calabi-Yau property.
	\item Actions of Hopf algebras and invariant theory.
	\item Representations of Hopf algebras, tensor and fusion categories.
\end{itemize}
\end{abstract}

\section{January 7, 2019}
If you don't know what a symmetric tensor category is, today is going to be a three 
star day. Max is 5.

\subsection{Overview}
We are shooting to understand two conjectures:

\begin{conj}[Etingof-Ostrik `04]
	If $A$ is a finite dimentional Hopf algebra, then 
	\[\bigoplus_{i\ge 0}\Ext_A^i(_Ak, _Ak)\]
	is Noetherian.
\end{conj}

\begin{conj}[Brown-Goodearl `98]
	If $A$ is a Noetherian Hopf algebra, then the injective dimension of $A_A$ is finite.
\end{conj}

These are both still open! In fact there is a meeting at Oberwolfach this March concerning
exactly these conjectures.

\subsection{Symmetric Tensor Categories}
We are going to be using the following notation throughout:
\begin{itemize}
	\item $k$ is a field
	\item $\Vectk$ is the category of $k$-vector spaces
	\begin{itemize}
		\item $\Vectk$ is closed under tensor products
		\item There is an element $k\in\Vectk$ such that
		\[k\otimes_k V\cong V\cong V\otimes_k k\]
		where the above isomorphisms are natural.
		\item $V\otimes_k W\cong W\otimes_k V$
	\end{itemize}
	\item An algebra is an object in $\Vectk$.
\end{itemize}

\begin{defn}\label{def-alg}
	$V\in\Vectk$ is called an \textbf{algebra object} if there are two morphisms
	\begin{enumerate}
		\item $m:V\otimes V\to V$
		\item $u:k\to V$
	\end{enumerate}
	such that the diagrams in figure \ref{fig-alg} commute.

\end{defn}
\begin{figure}\label{fig-alg}
	\begin{tikzcd}
		V\otimes V\otimes V\ar[r,"\id_V\otimes m"]\ar[d,"m\otimes \id_V"] & V\otimes V\ar[d,"m"]\\
		V\otimes V \ar[r,"m"] & V
	\end{tikzcd}
	\begin{tikzcd}
		k\otimes V\ar[r,"\sim"]\ar[rd,swap,"u\otimes\id_V"] & V & V\otimes k\ar[l,"\sim",swap]\ar[dl, "\id_V\otimes u"]\\
		& V\otimes V \ar[u,"m"] &
	\end{tikzcd}
	\caption{Diagrams for definition \ref{def-alg}.}
\end{figure}

\begin{lem}
	$V\in\Vectk$ is an algebra object iff $V$ is an algebra over $k$.
\end{lem}

\begin{lem}
	If $C$ is a symmetric tensor category, so is $C^{op}$.
\end{lem}

Then the natural thing to ask is: what is an algebra object in this opposite category?
\begin{defn}
	A \textbf{coalgebra object} in $C$ is an algebra object in $C^{op}$. Here we have 
	comultiplication $\Delta$ and counit $\varepsilon$.
\end{defn}
\begin{rmk}
	Naturally you could go about defining this from first principles and drawing the diagrams
	in figure \ref{fig-alg} with the arrows reversed, but we are probably mature enough
	to do without that (saving my fingers from repetitive strain injury in the process.)
\end{rmk}

\begin{lem}
	$\Algk$, defined as the category of algebra objects in $\Vectk$, is a symmetric tensor category. Furthermore
	$\Coalgk$, the category of coalgebra objects in $\Vectk$, is a symmetric tensor category.
\end{lem}

\begin{lem}
	The following are equivalent:
	\begin{enumerate}
		\item $V$ is an algebra object in $\Coalgk$
		\item $V$ is a coalgebra object in $\Algk$
		\item There are morphisms $m,u,\Delta,\varepsilon$ such that
		\begin{itemize}
			\item $(V,m,u)$ is an algebra
			\item $(V,\Delta,\varepsilon)$ is a coalgebra
			\item Equivalently:
			\begin{itemize}
				\item $m$ and $u$ are coalgebra morphisms
				\item $\Delta$ and $\varepsilon$ are algebra morphisms.
			\end{itemize}
		\end{itemize}
	\end{enumerate}
\end{lem}
\begin{prf}
	The nice thing here is that the $(a)\Leftrightarrow (c)$ without the last condition. A similar fact holds
	for $(b)$ except the second-to-last. The last thing to do is to prove the last two conditions are equivalent.
\end{prf}
\begin{prob}
	Fill in the details for the proof above.
\end{prob}
\begin{sol}
	Assume that $(V,m,u,\Delta,\varepsilon)$ is an algebra and coalgebra and further that
	$m$ and $u$ are coalgebra morphisms. That means in particular that the diagrams in figure \ref{fig-coalg-mor} commute.
	\begin{figure}[h]\label{fig-coalg-mor}
		\centering
		\begin{tikzcd}
			V\otimes V \ar[d,swap,"\Delta\otimes\Delta"]\ar[r,"m"] & V\ar[r,"\Delta"] & V\otimes V\\
			V^{\otimes 4}\ar[rr,"T_{2,3}"] & & V^{\otimes 4}\ar[u,swap,"m\otimes m"]
		\end{tikzcd}
		\begin{tikzcd}
			k\ar[r,"\Delta"]\ar[d,"u"] & k\otimes k\ar[d,"u\otimes u"]\\
			V\ar[r,"\Delta"] & V\otimes V
		\end{tikzcd}
		\label{fig-coalg-mor}
		\caption{$m$ and $u$ are coalgebra morphisms.}
	\end{figure}

	We are looking to prove that $\Delta$ and $\varepsilon$ are algebra morphisms, or that 
	the diagrams in figure \ref{fig-coalg-mor}  commute.
	\begin{figure}[h]
		\centering
		\begin{tikzcd}
			V\otimes V\ar[r,"m"]\ar[d,"\Delta\otimes \Delta"] & V\ar[d,"\Delta"]\\
			V^{\otimes 4}\ar[r,"m"] & V\otimes V
		\end{tikzcd}
		\begin{tikzcd}
			V\otimes V\ar[r,"\varepsilon\otimes\varepsilon"]\ar[d,"m"] & k\otimes k\ar[d,"m"]\\
			V\ar[r,"\varepsilon"] & k
		\end{tikzcd}
		\label{fig-alg-mor}
		\caption{$\Delta$ and $\varepsilon$ are algebra morphisms.}
	\end{figure}

	From here it's actually a bit boring because it's kinda just a definition/notation game.
	It boils down to the fact that the (co)multiplication on $V\otimes V$ has a twist that 
	exactly lines up so that each square is saying the same thing.
\end{sol}

\begin{defn}
	$V$ is called a \textbf{bialgebra object} if $V$ is an algebra object in $\Coalgk$.
\end{defn}

\begin{prob}
	\begin{enumerate}
		\item Suppose that $\ch k\ne 2$. Classify all bialgebras of $\dim 2.$
		\item Do the same for $\ch k = 2.$
	\end{enumerate}
\end{prob}
\begin{sol}
	\subsubsection*{Part (a)}
	Consider $\varepsilon:V\to k$ and consider $\ker\varepsilon\lhd V$. By rank-nullity, $\dim\ker\varepsilon= 1$, so
	$\ker\varepsilon=kx$ for some $x\in V$. Therefore $x^2=cx$ for some $c$, and if $c=0$, then (as an algebra) $V\cong k[x]/(x^2)$.
	Otherwise consider $y=\frac{x}{c}$. In this case $y^2=\frac{x^2}{c^2}=\frac{x}{c}=y$, and $V\cong k[x]/(x^2-x)$.

	Notice that in either case $\varepsilon(x)=0$, so let
	\[\Delta(x)=a(1\otimes 1)+b(1\otimes x)+c(x\otimes 1) + d(x\otimes x)\]
	and using that $\varepsilon\otimes \id\circ\Delta=\id\otimes\varepsilon\circ\Delta$
	and that each should be (essentially) the identity (this is just the diagram we saw before),
	we get $a=0$ and $b=c=1$. Thus the coalgebra structure of any Hopf algebra is given by 
	\[\varepsilon(x)=0,\quad \Delta(x)=1\otimes x+x\otimes 1+d(x\otimes x).\]

	Consider first the case when $x^2=0$. Then since comultiplication will be an algebra morphism,
	\[0=\Delta(x^2)=\Delta(x)^2=1\otimes x^2+x^2\otimes 1+d^2(x^2\otimes x^2)+2(x\otimes x)+2d(x\otimes x^2)+2d(x^2\otimes x)\]
	and since $x^2=0$, 
	\[0=2(x\otimes x).\]
	But $x\otimes x$ is a basis element of $V\otimes V$, so $V$ can only have this algebra structure
	when $\ch k=2$. We will return to this in the next part.

	So then $x^2=x$ and using the computation above,
	\[1\otimes x+x\otimes 1+d(x\otimes x)=\Delta(x)=\Delta(x^2)=1\otimes x+x\otimes 1+(d^2+4d+2)(x\otimes x)\]
	so
	\[(d^2+3d+2)(x\otimes x)=0\quad\Rightarrow\quad d^2+3d+2=(d+2)(d+1)=0\]
	and so either $d=-1$ or $d=-2$.

	One can verify that $\Delta$ is coassociative, so we can conclude that when $\ch k\ne 2$, 
	there are precisely two Hopf algebra structures with algebra structure $k[x]/(x^2-x)$ and
	comultiplication either
	\[\Delta(x)=1\otimes x+x\otimes 1-x\otimes x\quad\text{or}\quad \Delta(x)=1\otimes x+x\otimes 1-2(x\otimes x)\]

	\subsection*{(b)}
	Now assume that $\ch k=2$ and that $V\cong k[x]/(x^2-x)$ as an algebra. Then using the analysis above, 
	we see that we can choose comultiplication either
	\[\Delta(x)=x\otimes 1+1\otimes x\quad\text{or}\quad \Delta(x)=1\otimes x+x\otimes 1+x\otimes x.\]

	If instead $V\cong k[x]/(x^2)$, then \textit{any} value of $d$ will suffice, so there are a full $k$'s worth
	of Hopf algebra structures that can appear.
\end{sol}

\section{January 9, 2019}
Today we are going to rely heavily on Sweedler notation. :) Notice that if we are looking at actual objects in the 
diagram for coassociativity, we get
\begin{figure}[h]
	\centering
	\begin{tikzcd}
		v\ar[r]\ar[d] & \sum v_{(1)}\otimes v_{(2)}\ar[d]\\
		\sum v_{(1)}\otimes v_{(2)}\ar[r] & \sum (v_{(1)})_{(1)}\otimes (v_{(1)})_{(2)}\otimes v_{(2)}=\sum v_{(1)}\otimes (v_{(2)})_{(1)}\otimes (v_{(2)})_{(2)}
	\end{tikzcd}
	\caption{Coassociativity on elements in Sweedler notation}
\end{figure}

\begin{ex}
	Let $G$ be a group and $kG$ be the group algebra. The algebra structure arises as normal
	where $g\cdot h$ comes from the structure on $G$. Then $\Delta(g)=g\otimes g$ and this extends linearly.

	But then if you consider $\Delta(\sum c_gg)$, notice that by the nature of tensors this is not unique!
	So we will just write
	\[\Delta\left(\sum c_gg\right)=\sum_Gc_g(g\otimes g)=\sum v_{(1)}\otimes v_{(2)}\]
\end{ex}

\subsection{Algebra structure on $V\otimes V$}
We said earlier on that $\Algk$ is a symmetric \textit{tensor} category. But how do we
define the multiplication on the tensor product?

Well it all comes from the twist! We define
\[m_{V\otimes W}=(m_V\otimes m_W)\circ(\id_V\otimes \tau_{2,3}\otimes \id_W)\]
where $\tau_{2,3}$ is the twist morphism.

More simply, $u_{V\otimes W}:k=k\otimes k\to V\otimes V$ simply defined by $u_V\otimes u_W$.

So then when we say that $\Delta$ is an algebra morphism, we are saying that for all $v,w\in V$
\[\Delta(vw)=\sum(vw)_{(1)}\otimes(vw)_{(2)}=(\sum v_{(1)}\otimes v_{(2)})(\sum w_{(1)}\otimes w_{(2)})=\sum v_{(1)}w_{(1)}\otimes v_{(2)}w_{(2)}\]

\subsection{Hopf Algebras}
Already to the good stuff!
\begin{defn}
	$V\in\Vectk$ is a \textbf{Hopf algebra} if $V$ is a bialgebra together with an \textbf{antipode}
	$S:V\to V$ satisfying
	\[(S,\id_V)\circ \Delta=\varepsilon=(\id_V,S)\circ\Delta\]
\end{defn}
\begin{conj}
	If $V\in\Vectk$ is a Noetherian Hopf algebra, then $S$ is bijective.
\end{conj}

\subsection{History and Motivation}
Hopf himself was a topologist, so this is the first context in which it arose. In the 1940's,
he began studying Hopf algebras over $\Z_2$ graded $k$ vector spaces. For instance, the 
cohomology ring of topological space $X$ with coefficients in $k$.

Later, in combinatorics, they ended popping up. Looking at rings of symmetric functions and other places
gave some interesting examples.

Then in group theory you can define a functor from groups to Hopf algebras by $F(G)=kG$
with the diagonal map. The antipode is just the inverse.

Then with Lie algebras, you can look at $\mathcal{U}(L)$, the universal enveloping algebra is a Hopf algebra.

Finally with algebraic groups (yay!) we take an algebraic group $G$ and consider the ring of functions on it, 
which is again a Hopf algebra.

\brk

Some ``cousins'' of Hopf algebras: quasi, weak, multiplier, ribbon, quasi-triangular, etc Hopf algebras. 
Each has slightly different base category or restrictions.

\section{January 11, 2019}
The plan for today is to talk about:
\begin{itemize}
	\item Convolution Algebras
	\item Antipodes
	\item Duality
	\item (Co-)Modules
\end{itemize}

\subsection{Convolution Algebras}
 Let $\scrT$ be a symmetric tensor category. We can usually think of $\scrT=\Vectk$, but 
 there is a problem since $\Vectk$ is equivalent to the category of Hopf algebras over $k$,
 while this is not generally true.

 We also need that $\scrT$ is $k$-linear (that is, enriched as a category over $k$). This
 means that $\Hom_\scrT (A,B)\in\Vectk$.

 \begin{thm}
	Let $\scrT$ be as above. Then $\Hom_\scrT(C,A)$ is an algebra and $\Hom_\scrT(-,-):(\Coalg_\scrT)^{op}\times\Alg_\scrT\to\Alg_k$
	is a functor.
 \end{thm}
 \begin{prf}
	Let $A$ be an algebra object in $\scrT$ and $C$ be a coalgebra object in $\scrT$. Then $1_{\Hom} :=u_A\circ\varepsilon_C:C\to 1_\scrT$ and define
	\[f\ast g := m_A(f\otimes g)\Delta_C:C\to A.\]

	Then using Lemma \ref{lem-func} and the fact that $A$ and $C$ are (co)algebra objects,
	we can see that the product $\ast$ satisfies the axioms required.

	Note that actually
 \end{prf}
 \begin{lem}
	\begin{enumerate}
		\item $1_{\Hom}\ast f = m_A(u\otimes 1)(1\otimes f)(\varepsilon\otimes 1)\Delta.$
		\item $f\ast 1_{\Hom}=m_A(1\otimes u)(f\otimes 1)(1\otimes \varepsilon)\Delta$
		\item $(f\ast g)\ast h=m_A(m_A\otimes 1)((f\otimes g)\otimes h)(\Delta_C\otimes 1)\Delta_C$
		\item $f\ast(g\ast h)=m_A(1\otimes m_A)(f\otimes(g\otimes h))(1\otimes \Delta)\Delta$
	\end{enumerate}
	\label{lem-func}
 \end{lem}

 \begin{prf}
	\subsubsection*{(a)}
	\begin{align*}
		1_{\Hom}\ast f&= m_A(1_{\Hom}\otimes f)\Delta_C\\
		&=m_A(u_A\circ\varepsilon\otimes f)\Delta_C\\
		&=m_A(u\otimes 1)(\varepsilon\otimes 1)(1\otimes f)\Delta
	\end{align*}
	\subsubsection*{(b)}
	Same as (a), essentially.
	\subsubsection*{(c) and (d)}
	\begin{align*}
		(f\ast g)\ast h&= m_A((f\ast g)\otimes h)\Delta\\
		&=m_A((m_A(f\otimes g)\Delta)\otimes h)\Delta\\
		&=m_A(m_A\otimes 1)((f\otimes g)\otimes h)(\Delta\otimes 1)\Delta
	\end{align*}
	and the other is analogous.
 \end{prf}

 \begin{defn}
	$V\in\scrT$ is a \textbf{Hopf algebra object} if:
	\begin{itemize}
		\item $V$ is a bialgebra object in $\scrT$ and
		\item There is a map $S:V\to V$ that is $(\id_V)^{-1}$ with respect to $\ast$.
	\end{itemize}
 \end{defn}
 \begin{rmk}
	Notice here that $\id_V\in\Hom_\scrT(V,V)$, the identity map in $\scrT$. We are \textit{not}
	taking about $1_{\Hom}=u\circ\varepsilon$.

	Also, we call $S$ an \textbf{antipode.}
 \end{rmk}

 \subsection{Duality}
 Notice that when $C=1_\scrT$ (that is the tensor identity), $\Hom_\scrT(1_\scrT,-):\Alg_\scrT\to\Algk$
 is a functor. Same for the dual from $\Coalg_\scrT$. This second one gives us a chance to talk about duality.

 \begin{lem}
	Let $\scrT$ be the category of finite dimensional vector spaces over $k$.
	Then $(-)^*:\scrT\to\scrT^{op}$ is an equivalence.
 \end{lem}
 This uses $(V\otimes W)^*=W^*\otimes V^*$.
 \begin{cor}
	$V$ is an algebra over $k$ $\Leftrightarrow$ $V^*$ is a coalgebra over $k$.
	And vice versa.
 \end{cor}

 Recall $S=(\id_V)^{-1}$. Thus
 \[S\ast \id_V=1_{\Hom}=\id_V\ast S\]
 The diagram we have here is
 \begin{figure}[h]
	 \centering
	 \begin{tikzcd}
		V\otimes V\ar[rr,"S\otimes \id_V"] & & V\otimes V\ar[d,"m_V"]\\
		V\ar[u,"\Delta"]\ar[d,"\Delta"]\ar[r,"\varepsilon_V"] & k \ar[r,"u_V"] & V\\
		V\otimes V\ar[rr,"\id_V\otimes S"] & & V\otimes V\ar[u,"m_V"]
	 \end{tikzcd}
 \end{figure}

 \subsection*{Modules/Comodules}
 \begin{defn}
	Let $A$ be an algebra object in $\scrT$. A \textbf{left $A$ module} is $M\in\scrT$
	with a morphism
	\[m_M:A\otimes M\to M\]
	such that the diagrams in Figure \ref{fig-mod} commute.
\end{defn}
\begin{figure}[h]\label{fig-mod}
	\centering
	\begin{tikzcd}
		A\otimes A\otimes M\ar[r,"m_A\otimes 1"]\ar[d,"1\otimes m_M"] & A\otimes M\ar[d,"m_M"]\\
		A\otimes M\ar[r,"m_M"] & M
	\end{tikzcd}
	\begin{tikzcd}
		A\otimes M\ar[r,"m_M"] & M\\
		1_\scrT\otimes M\ar[ur,"\sim"]\ar[u,"u_A\otimes 1"] &
	\end{tikzcd}
	\caption{Module diagrams}
\end{figure}
\begin{rmk}
	Note that we don't necessarily need that $M$ lie in $\scrT$. We could instead just 
	rely on an algebra homomorphism $\varphi:A\to \Hom_{\scrT}(M,M)$ and proceed as usual.
\end{rmk}

\section{January 14, 2019}
Today we are talking about Hopf modules and later in the week we will see the fundamental
theorem of Hopf modules as well as a neat result.

Here is the ``example of the day.''
\begin{ex}
	Consider $k[x]$ with maps $\Delta(x^n)=\sum_0^nx^i\otimes x^{n-i}$ and $\varepsilon(x^n)=\delta_{1,n}$. 
	$S(x^n)=(-x)^n$. Then notice that $\Delta(x)=1\otimes x+x\otimes 1$, so $\delta(x^n)=(\delta x)^n$.

	James doesn't want to do the rest of the computations, but they can be done. :)
\end{ex}

\begin{prob}\label{prob-4.1}
	(**) Working with a Hopf algebra $V$, consider the convolution algebra $\Hom(V\otimes V,V)$
	\begin{enumerate}
		\item $(S\circ m)\ast m=1_{\Hom(V\otimes V,V)}=m\ast(m\circ s\otimes s\circ\tau)$ where $s$ is the antipode in $V$.
		\item $(\Delta\circ S)\ast \Delta=1_{\Hom(V,V\otimes V)}=\Delta\ast((s\otimes s)\circ\tau\circ \Delta)$
	\end{enumerate}
\end{prob}
\begin{sol}
	\begin{align*}
		(s\circ m)\ast m
	\end{align*}
\end{sol}
\begin{prob}
	Classify all Frobenius (to be defined) Hopf algebras of dimension 3.
\end{prob}

\subsection{Returning to (co)modules}\label{subsec-comod}
\begin{lem}
	Let $\scrT=\Vectk$, and $V$ and algebra over $k$. The following are equivalent:
	\begin{enumerate}
		\item $M$ is a left $V$-module (see last lecture).
		\item There is an action of $A$ on $M$ such that $(ab)\cdot m=a\cdot (b\cdot m)$ and $1\cdot m=m$.
		\item There is an algebra morphism $\varphi:V\to\Hom_k(M,M)$
	\end{enumerate}
\end{lem}

\begin{defn}
	Let $V$ be a coalgebra over $k$. We say $M$ is a left \textbf{comodule} if there is a map
	\[\rho_M:M\to V\otimes M\]
	such that the diagrams in figure \ref{fig-com} commute.
\end{defn}
\begin{figure}[h]\label{fig-com}
	\begin{tikzcd}
		M\ar[r,"\rho"]\ar[d,"\rho"] & V\otimes M\ar[d,"\id_V\otimes\rho"]\\
		V\otimes M\ar[r,"\Delta_V\otimes\id_M"] &V\otimes V\otimes M
	\end{tikzcd}
	\begin{tikzcd}
		M\ar[r,"\rho"]\ar[d,"\cong"] & V\otimes M\ar[dl, "\varepsilon\otimes\id_M"]\\
		k\otimes M
	\end{tikzcd}	
	\caption{Diagrams for the definition of a comodule.}
\end{figure}

\begin{rmk}
	Notationally speaking, we write $_V\mathcal{M}$ and $^V\mathcal{M}$ for the categories of 
	left modules and comodules. Similar for right boyes.
\end{rmk}

\subsection{Tensors of $V$-modules}
Say $V$ is a bialgebra (or Hopf if you prefer, but it's not necessary). Then we get the following:
\begin{lem}
	Let $M$ and $N$ be two right $V$-modules. Then
	\begin{enumerate}
		\item $M\otimes N$ is a right $V$-module.
		\item If $V$ is cocommutative, then $M\otimes N\cong N\otimes M$ as right $V$-modules.
		\item $\mathcal{M}_V$ is a tensor (monoidal) category.
	\end{enumerate}
\end{lem}
\begin{prf}
	For (a), use the fact that $V\otimes V$ acts on $M\otimes N$ in a natural way. Then you get
	a map $V\otimes V\to \Hom_k(M\otimes N,M\otimes N)^{op}$ (why opposite?) and by precomposing
	with $\Delta$ to show they are $V$-modules. This establishes (b).

	Finally (c) follows because $\Vectk$ is a tensor category.
\end{prf}

\subsection{Hopf Modules}\label{subsec-hopfmod}
\begin{defn}
	Let $V$ be a bialgebra over $k$. We say that $M$ is a $\binom{r}{r}$ Hopf $V$-module if
	\begin{enumerate}
		\item $(M,m_M)$ is a right $V$-module
		\item $(M,\rho_M)$ is a right $V$ comodule.
		\item $\rho_M$ is a right $V$-module map.
		\begin{itemize}
			\item Equivalently, $m_M$ is a right $V$-comodule map.
		\end{itemize}
	\end{enumerate}
\end{defn}
\begin{rmk}
	The fact that these two last conditions are equivalent are perhaps not immediately obvious but we
	have been assured it is true. :)
\end{rmk}

\section{January 16, 2019}
\begin{thm}[Larson-Sweedler `69]\label{thm-LS69}
	Let $V$ be a Hopf algebra over $k$. Then the following categories are equivalent:
	\[_V^V\mathcal{M}\cong \Vectk\cong \mathcal{M}_V^V\]
\end{thm}

\begin{rmk}
	A natural question that one may ask is how to extend this theorem to the enriched setting.
\end{rmk}

\begin{lem}
	$S$ is an anti-homomorphism of an algebra $V$ and an anti-homomorphism of the coalgebra $V$.
\end{lem}
\begin{prf}
	Use the facts proved in problem~\ref{prob-4.1}.
\end{prf}

\begin{prob}
	If $\ch k=0$ any Hopf quotient of $k[x]$ is either $k$ or $k[x]$ itself.
\end{prob}

\begin{prob}
	Suppose $k$ has positive characteristic. Construct two different Hopf algebra quotients of $k[x]$ of dimension $p=\ch k$.
\end{prob}

\begin{ex}
	$\ch k=p>0$. Let $V=k[x]/(x^p)$ as an algebra and define $\Delta(x)=x\otimes 1+1\otimes x$,
	$\varepsilon(x)=0$ and $S(x)=x$. Then notice
	\[\Delta(x^p)=\sum\binom{p}{i}x^i\otimes x^{p-i}=x^p\otimes 1+1\otimes x^p.\]
\end{ex}

\subsection{Hopf Modules Once More}
Recall the definitions from section~\ref{subsec-comod} and section~\ref{subsec-hopfmod}.

\begin{rmk}
	An equivalent (more category-theoretical) definition of a Hopf module is a $V$-comodule object in 
	the category of $V$-modules. Also can dualize everything.
\end{rmk}

\begin{ex}
	Consider $V\in _V^V\mathcal{M}$. THen you can define $\rho=\Delta$ and check it satisfies all the requirements.
\end{ex}
\begin{defn}
	An $\binom{l}{l}$ Hopf $V$-module $M$ is called \textbf{trivial} if $M\cong V\otimes M_0$ for some $M_0\in\Vectk.$
\end{defn}
\begin{rmk}
	Basically you can think of this being trivial since we can \textit{always} define a module in this way
	where the entire module structure is inherited from the structure of $V$ (that is, irrespective of $M_0$).
\end{rmk}
\begin{thm-prime}{thm-LS69}%[Fundamental Thm of Hopf Modules (L-S `69)]
	Suppose that $V$ is a Hopf algebra. Then
	\begin{enumerate}
		\item Every $\binom{l}{l}$ Hopf $V$-module is trivial.
		\item Let $M\in _V^V\mathcal{M}$. Then 
		\[M\xrightarrow{\sim} V\otimes M^{Cov}\]
		where
		\[M^{Cov}:=\{m\in M|\rho(m)=1\otimes m\}.\]
		\item $_V^V\mathcal{M}\cong \Vectk$.
	\end{enumerate}
\end{thm-prime}
\begin{rmk}
	This is actually just a reformulation of theorem~\ref{thm-LS69} in less compact (but ultimately more usable) notation.
\end{rmk}
\begin{prf}
	Some identites that will be useful:
	\begin{align*}
		\Delta(v)=\sum v_1\otimes v_2\\
		\sum S(v_1)\otimes v_2=\varepsilon(v)=\sum v_1S(v_2)\\
		\rho(m)=\sum m_{-1}\otimes m_1\\
		\sum v_1\otimes (v_2)_1\otimes (v_2)_2=\sum (v_1)_1\otimes (v_1)_2\otimes v_2=\sum v_1\otimes v_2\otimes v_3\\
		\sum (m_{-1})_1\otimes (m_{-1})_2\otimes m_0=\sum m_{-1}\otimes (m_0)_{-1}\otimes (m_0)_0=\sum m_{-2}\otimes m_{-1}\otimes m_0
	\end{align*}

	Then we defin $\phi:M\to M^{Cov}$ by 
	\[\phi(x)=\sum S(x_{-1})x_0\in M\]
	We claim first that $\phi(x)\in M^{Cov}$ -- namely $\rho(\phi(x))=1\otimes\phi(x)$. To see this, compute
	\begin{align*}
		\rho(\phi(x)) &= \sum (\phi(x))_{-1}\otimes (\phi(x))_0\\
		&= (S(x_{-1})x_0)_{-1}\otimes (S(x_{-1})x_0)\\
		&= \sum\Delta(S(x_{-1}))\rho(x_0)\\
		&= \sum (S\otimes S)\circ \tau(\Delta(x_{-2}))\cdot [x_{-1}\otimes x_0]\\
		&=\sum[S(x_{-2})\otimes S(x_{-3})][x_{-1}\otimes x_0]\\
		&= \text{my fingers are aching...}\\
		&= 1\otimes\phi(x)
	\end{align*}

	Now define $F:M\to V\otimes M^{Cov}$ by $F=(\id\otimes\phi)\circ\rho$ and $G:V\otimes M^{Cov}$ 
	to be the map taking $v\otimes m$ to $vm$. The next claim is that $GF$ and $FG$ are the identity.
	You can see this by similarly pushing around notation.

	Finally the last claim will be seen on Friday.
\end{prf}
\section{January 18, 2019}
Recall the Fundamental Theorem of Hopf Modules: $_V^V\mathcal{M}\simeq \Vectk$ or equivalently that
every (not necessarily finite dimensional) Hopf module is trivial: $M\cong V\otimes M^{Cov}$.

\section{Frobenius Algebras}
Today we will discuss the result that
\begin{thm}[Larson-Sweedler, `69]\label{LS-69-2}
	Every finite dimensional Hopf algebra is Frobenius.
\end{thm}

There are some nice categorical equivalences: finite dimensional Frobenius algebras over $k$ is
equivalent to 2-D quantum field theories, or equivalently the symmetric functor category from the 2-D cobordism category to $\Vectk$.

Whoa.

\begin{ex}
	$V=(kG)^*=\oplus_{g\in G}k\delta g$ where $G$ is a finite group. Define the algebra structure with
	\[1_V=\sum_{g\in G}\delta_g\]
	and $\delta_g\delta_h=\delta_g$ when $g=h$ and 0 otherwise.

	The coalgebra structure is $\varepsilon(\delta_g)=\delta_{g,1_G}$ where on the right it is the Kroneker delta.
	Furthermore $\Delta(\delta_g)=\sum_{h\in G}\delta_h\otimes\delta_{h^{-1}g}$ and $S(\delta_g)=\delta_{g^{-1}}$.
\end{ex}

\begin{prob}
	Prove that $kS_3\cong(kS_3)^*$.
\end{prob}
\begin{prob}
	Prove that $k\Z_3\cong(k\Z_3)^*$ when $\ch k\ne 3$.
\end{prob}

\begin{lem}
	Let $A$ and $B$ be algebras over $k$.
	\begin{enumerate}
		\item If $M\in_A\mathcal{M}_B$, $N\in \mathcal{M}_B$, then $\Hom_{\mathcal{M}_B}(_AM,N)\in\mathcal{M}_A$.
		\item Take $B=k$, $M\in _A\mathcal{M}$. Then
		\[M^*:=\Hom_k(M,k)\in\mathcal{M}_A.\]
		\item $(-)^*:_A\mathcal{M}_{f.d.}\to (\mathcal{M}_A)_{f.d.}^{op}$ is an equivalence of categories. This is called the \textbf{reflection principle.}
	\end{enumerate}
\end{lem}
\begin{prf}
	This is a known fact. (b) follows quickly from (a) and (c) involves extending the dual map to maps of left $A$-modules.
\end{prf}
By the principle of duality, we have:
\begin{lem}
	If $C$ and $d$ are finite dimensional coalgebras over $k$, then
	\begin{enumerate}
		\item if $M\in ^C\mathcal{M}^D_{f.d.}$ and $N\in \mathcal{M}^D_{f.d.}$ then $\Hom_{\mathcal{M}^D}(M,N)\in\mathcal{M}^C_{f.d.}$.
		\item If $M\in ^C\mathcal{M}_{f.d.}$ then $\Hom_k(M,k)\in\mathcal{M}^C$
		\item $(-)^*:^C\mathcal{M}_{f.d.}\to(\mathcal{M}_{f.d.}^C)^{op}$ is an equivalence of categories.
	\end{enumerate}
\end{lem}

\begin{defn}
	A finite dimensional algebra $A$ is called \textbf{Frobenius} if one of the following (equivalent)
	conditions hold:
	\begin{enumerate}
		\item $(A)^*_A\cong A_A$
		\item $_A(A)^*\cong _AA$
	\end{enumerate}
\end{defn}
\begin{rmk}
	Note that the equivalence of these two things follow due to the equivalence of categories we get
	from the two lemmas above.
\end{rmk}
We actually don't need the following lemma, but James is a fan so we're going to write it down.
\begin{lem}\label{lem-not-needed}
	Let $A$ be a finite dimensional algebra over $k$. Then $M$ is a finite dimensional left $A$-module if
	and only if $M$ is a finite dimensional right $A^*$-comodule.
\end{lem}
\begin{rmk}
	The reason we get the module to comodule switch is (partly?) due to the fact that in our
	monoidal category $(V\otimes W)^*=W^*\otimes V^*.$
\end{rmk}

\begin{lem}
	Let $V$ be a finite dimensional Hopf algebra and $M$ be a finite dimensional $\binom{l}{l}$ Hopf $V$-module.
	Then $M^*$ is a finite dimensional $\binom{r}{r}$ Hopf $V$-module.
\end{lem}
\begin{prf}
	$M\in {_V^V\mathcal{M}}$ if and only if $M\in_V\mathcal{M}\cap {^V\mathcal{M}}$ and the two structures are compatible.

	The fact that $M^*\in \mathcal{M}_V\cap\mathcal{M}^V$ is straightforward and checking the compatibility is easy to check.
\end{prf}

Now we finally prove that theorem~\ref{LS-69-2}:
\begin{prf}
	First of all $V\in{^V_V\mathcal{M}}$ so by the lemma $V^*\in\mathcal{M}_V^V$. By the fundamental theorem of Hopf modules,
	\[V^*\cong V\otimes(V^*)^{Cov}\cong V\]
	as Hopf modules. Thus $V_V^*\cong V_V$ in particular.
\end{prf}
\begin{rmk}
	The second isomorphism above comes from the fact that $(V^*)^{Cov}$ is one-dimensional, so is the trivial module. :)
\end{rmk}
\begin{cor}
	If $A$ is \textit{not} a Forbenius algebra, then there is \textit{no} Hopf algebra structure on $A$.
\end{cor}
\begin{ex}
	Some examples of algebras with(out) this property.
	\begin{enumerate}
		\item $\C[x]/(x^n)$ is finite dimensional and Frobenius, but if $n\ge 2$, there is \textbf{no Hopf structure.}
		\item $\C[x,y]/(x^2,y^2,xy)=\C1+\C x+\C y$. This has the property that it is local (unique maximal submodule over itself)
		but $A^*$ has two maximal submodules, so $A\not\cong A^*$, so not Frobenius.
	\end{enumerate}
\end{ex}
Finally we conclude with a proof of lemma~\ref{lem-not-needed}:
\begin{prf}
	Say that $\{a_i\}$ is a basis for $A$ and $\{a_i^*\}$ is a basis for $A^*$ ($a_i^*(a_j)=\delta_{ij}$).

	Define $\rho(m)=\sum(a_i\cdot m)\otimes a_i^*$. A lemma (use change of basis matrices) shows that
	$\sum a_i\otimes a_i^*$ is independent of choice of basis. Another lemma says (scalar) multiplication is associative
	if and only if $\rho$ is coassociative. This gets us the forward direction.

	For the reverse direction, say $\rho(m)=\sum m_0\otimes m_{-1}.$ then define $a\cdot m=\sum m_1(a)m_0$.
\end{prf}

\section{January 23, 2019}
Today we are interested in studying the representation category $_V\mathcal{M}$ for any Hopf algebra $V$. It will end up
(and this is a bit cryptic for now) that $k$ controls everything here.
\section{Represetnations and Modules}
Notice that for any Hopf algebra $V$ we have the exact sequence:
\[0\to \ker\varepsilon\to V\xrightarrow{\varepsilon} k\to 0\]
\begin{defn}
	The trivial $V$ modules is $V/\ker\varepsilon=k\in _V\mathcal{M}_V$.
\end{defn}
\begin{ex}
	$V=kG$ where $G$ is a group, $\varepsilon(g)=1$ for all $g\in G$. Then $\ker\varepsilon=\oplus k(g-1)$ and so
	\[k=V/\oplus k(g-1).\]
\end{ex}
\begin{ex}
	$V=k[x]$ where $\Delta(x)=1\otimes x+x\otimes 1$ and $\varepsilon(x)=0$. Then $\ker\varepsilon=\oplus_{n\ge 1}kx^n$, whence
	\[V/\ker\varepsilon=k[x]/(x).\]
\end{ex}
\begin{ex}
	Let $V=(kG)^*=\oplus_G k\delta_g$ and $\varepsilon(\delta_g)=\delta_{1,1_G}$. THen $\ker\varepsilon=\oplus_{g\ne 1_G}k\delta_g$ and
	\[V/\ker\varepsilon=k\delta_1\]
\end{ex}
The following uses $T_4$, the example of the day I missed due to my meeting.
\begin{ex}
	$T_4$: $\varepsilon(g)=1$, $\varepsilon(p)=0$. Then $\ker\varepsilon=k(g-1)\oplus kp\oplus kpg$ and
	\[V/\ker\varepsilon=k.\]
\end{ex}
\begin{lem}
	\begin{itemize}
		\item For all $x \in V$ and $a\in k$ (the trivial module), $xa=\varepsilon(x)a$.
		\item $k$ is the identity object in $_V\mathcal{M}$.
		\item If $V$ is cocommutative, then $_V\mathcal{M}$ is a symmetric tensor category.
	\end{itemize}
\end{lem}
\begin{prf}
	\[x\cdot a=(x-\varepsilon(x))a+\varepsilon(x)a=\varepsilon(x)a\]
	since $x-\varepsilon(x)\in\ker\varepsilon$.

	For the next part, consider that for every $M\in _V\mathcal{M}$, we have 
	\[M\otimes k\cong M\cong k\otimes M.\]
	But then for all $x\in V$, we have
	\begin{align*}
		\varphi[x\dot(m\otimes 1)]&=\varphi(x_1\cdot m\otimes x_2\cdot 1)\\
		&=\varphi(x_1\cdot m\otimes \varepsilon(x_2)\cdot 1)\\
		&=\varphi((x_1\varepsilon(x_2))m\otimes 1)\\
		&=\varphi(xm\otimes 1)=xm=x\varphi(m\otimes 1).
	\end{align*}
\end{prf}
\subsection{Integrals}
\begin{defn}
	An element $x\in V$ is called a \textbf{left integral} if $vx=\varepsilon(v)x$ for all $v\in V$.
	Similar for \textbf{right integral.}
\end{defn}
\begin{lem}
	The following are equivalent:
	\begin{itemize}
		\item $x\in V$ is a left integral.
		\item $kx\cong$ the trivial module.
		\item $1\mapsto x$ defines a left $V$-module morphism $k\to kx$.
	\end{itemize}
\end{lem}
\begin{lem}
	Let $\int_V^l$ denote the set of left integrals in $V$.
	\begin{itemize}
		\item $\int_V^l$ forms a vector subspace of $V$.
		\item $\int_V^l$ forms a left $V$ submodule of $V$. Thus it forms a left ideal of $V$.
		\item $\int_V^l$ forms a right ideal of $V$.
		\item $\int_V^l\cong\Hom_V(k,V)$ (left $V$-mods)
	\end{itemize}
\end{lem}
\begin{prf}
	The first three are relatively obvious. The last one comes from the map $\Phi:x\in\int_V^l\mapsto f_x:1\to x.$
\end{prf}
\begin{rmk}
	Unfortunately it ends up that the collection above ends up usually being zero. 
\end{rmk}
\begin{thm}
	Let $V$ be a finite dimensional Hopf algebra. Then $\dim\int_V^l=1$.
\end{thm}
\begin{prf}
	\begin{align*}
		\dim\int_V^l&= \dim\Hom_V(_Vk,_VV)\\
		&=\dim\Hom_V((V^*)_V,k_V^*)\\
		&=\dim\Hom_V(V_V,k)=\dim k=1
	\end{align*}
	Where we used above that $V$ was finite dimensional whence Frobenius to get $V^*\cong V$.
\end{prf}
\begin{ex}
	$V=kG$ for $G$ a finite group. Then $\int_V^l=\sum_G g$ and
	\[v\cdot(\sum_G g)=\varepsilon(v)(\sum_G g)=\int_V^r\]
	and when $v=h\in G$, we can compute it.
\end{ex}
\begin{ex}
	When $V=k[x]$, which is infinite dimensional, we get no integrals. This is because it is an 
	integral domain, so if
	\[x\cdot\int_V^l=\varepsilon(x)\int_V^l=0\cdot\int_V^l=0\]
	then this forces $\int_V^l=0$.
\end{ex}
\begin{ex}
	When $V=(kG)^*$ for $G$ a finite group, then 
	\[\int_V^l=\int_V^r=\delta_{1_G}\]
\end{ex}
\begin{ex}
	When $V=T_4=k1\oplus kg\oplus kp\oplus kgp$,
	\[\int_V^l=(g+1)p\ne\int_V^r=p(1+g)=(1-g)p\]
\end{ex}
\begin{ex}
	When $\ch k=p$ and $V=k[x]/(x^p)$, then 
	\[\int_V^l=\int_V^r=x^{p-1}.\]
\end{ex}

\section{January 28, 2019}
Today we will be mostly doing things in homological algebra. :)
\begin{ex}
	An example of an infinite dimension Hopf algebra is of the following: fix $q\in k^*$. Then 
	\[T_\infty=k<g, g^{-1}, p>/\langle gg^{-1}=g^{-1}g=1, gp=qpg\rangle\]
	and we can find $\{g^ip^j:i\in\Z, j\in\N\}$ is a $k$-linear basis.

	For the coalgebra structure, define $\Delta(g)=g\otimes g$ and $\varepsilon(g)=1$ and 
	$\Delta(p)=p\otimes 1+g\otimes p$ and $\varepsilon(p)=0$. Finally $S(g)=g^{-1}$ and $S(p)=q^{-1}pg^{-1}$.
\end{ex}

\subsection{Some nice results in Homological algebra}
\begin{prob}
	Prove that $\Q$ is an injective $\Z$-module.
\end{prob}
\begin{thm}[Auslander-Buchsbaum `59]
	Every local commutative algebra with finite global dimension is a UFD.
\end{thm}
Another result is about algebraic groups:
\begin{thm}
	If $\ch k=0$, then every Noetherian commutative Hopf algebra has finite global dimension
	and is a direct sum of integral domains, each of which is isomorphic.
\end{thm}
\begin{defn}[Semiprime]
	An ideal $I\lhd R$ is called \textbf{semiprime} if it is the intersection of (possibly infinitely many) primes.
\end{defn}
\begin{conj}
	Every Noetherian Hopf algebra of finite global dimension is semiprime (that is, 0 is semiprime in $V$).
\end{conj}

\begin{defn}[Connected Graded]
	An algebra $A$ is called \textbf{connected graded} if 
	\[A=k\oplus A_1\oplus A_2\oplus \cdots\]
	and $1_A=1_k\in k$, $A_iA_j\subseteq A_{i+j}$.

	The new part (connected) refers to the fact that $k$ is a summand.
\end{defn}
\begin{thm}
	Suppose that $\ch k=0$. Let $V$ be a Noetherian Hopf algebra that is connected graded as an algebra (says nothing about the coalgebra structure).
	Then the following hold:
	\begin{itemize}
		\item $V$ has finite global dimension.
		\item $V$ is a domain.
		\item $V$ is Artin-Schelter regular.
		\item $V$ is Auslander regular and Cohen-Macaulay.
		\item $V$ is Calabi-Yau
		\item $V$ is an iterated Ore extension.
	\end{itemize}
\end{thm}
That's a lot of word salad.

\subsection{Projective Modules}
\begin{defn}[Module-Theoretic Definitions]
	Let $M$ be a (left) $A$ module. 
	\begin{enumerate}
		\item $M$ is \textbf{free} if $M\cong \oplus_\mathcal{I} A$
		\item $M$ is \textbf{projective} if it is a direct summand of a free module.
	\end{enumerate}
\end{defn}
\begin{lem}
	$p\in{_A\mathscr{M}}$ is projective if and only if the diagram in figure~\ref{fig-proj2} commutes.
\end{lem}
\begin{figure}[h]
	\centering
	\begin{tikzcd}
		& P\ar[d,"f"]\ar[dl,dashed,"F",swap] &\\
		M\ar[r,"\pi"] & N\ar[r] & 0
	\end{tikzcd}
	\caption{The definition of a projective object.}
	\label{fig-proj2}
\end{figure}
\begin{rmk}
	This actually is just the definition for any Abelian category. This may generalize even further, but this is enough for us. :)
\end{rmk}

\begin{defn}[Injective Module]
	An \textbf{injective} $V$-module is a projective $V^{op}$ module.
\end{defn}

\section{January 30, 2019}
Today we're talking more about homological algebra. In particular, we'll learn about (or see again)
\begin{itemize}
	\item Complexes
	\item Projective resolutions and dimension
\end{itemize}
First the example of the day:
\begin{ex}
	Consider $GL_2(k)$, a group -- in fact, an algebraic group! Then 
	\[\mathcal{O}(GL_2)=k[x_{11},x_{12},x_{21},x_{22},det^{-1}]/\langle det^{-1}(x_{11}x_{22}-x_{21}x_{12}-1\rangle\]
	Let $X=(x_{ij})\in GL_2(k)$.

	Define the coalgebra structure via $\Delta(X)=X\otimes X$ and $\Delta(x_{ij})=\sum_1^2 x_{is}\otimes x_{sj}$,
	$\varepsilon(X)=I_2$, $\varepsilon(x_{ij})=\delta_{ij}$. Some computation (this is not trivial!) give us
	\[\Delta(det)=det\otimes det\quad\Rightarrow\quad \Delta(det^{-1})=det^{-1}\otimes det^{-1}\]
	Then we can find that $S(X)=X^{-1}$ and $S(det^{\pm})=det^{\mp}$.
\end{ex}
\begin{prob}
	Construct $\mathcal{O}(GL_n)$.
\end{prob}
\begin{prob}
	Construct $\mathcal{O}(G)$.
\end{prob}
\subsection{Homological stuff}
\begin{defn}[Complex]
	A \textbf{complex} of $A$-modules is a sequence of $A$-modules connected by homomorphisms
	\[\cdots\to M_2\xrightarrow{d_2}M_1\xrightarrow{d_1} M_0\to \cdots\]
	where $d_i\circ d_{i+1}=0$ for all $i$.
\end{defn}
Then we went over the standard definitions for:
\begin{itemize}
	\item Homology
	\item (Short) Exact sequences
	\item Projective resolutions
	\item Projective dimension
\end{itemize}
\begin{defn}[Global (Homological) Dimension]
	The (left) \textbf{global dimension} is
	\[\gldim A=\max\{\projdim M| M\in {_A\mathcal{M}}\}\]
\end{defn}
\begin{thm}[Hilbert Syzygy Theorem]
	\[\gldim k[x_1,\dots,x_n]=n\]
\end{thm}
\begin{cor}
	\[\gldim k= 0\]
\end{cor}
\begin{defn}
	The \textbf{finitistic global dimension} of $A$ is defined to be
	\[\findim A=\max\{\projdim M| M\in{_A\mathcal{M}}, M\text{ is f.g.},\projdim M <\infty\}\]
\end{defn}
\begin{conj}[Finitistic Dimension Conjecture]
	Let $A$ be a finitely generated algebra. Then $\findim A<\infty$.
\end{conj}
\section{February 1, 2019}
A really cool theorem:
\begin{thm}[Lorenz-Lorenz `95]
	Let $V$ be a Hopf algebra. Then $\gldim V=\projdim {_Vk}$.
\end{thm}
\begin{prf}
	It is enough to show that $\gldim V\le\projdim{_Vk}$. Or equivalently that
	\[\projdim M\le \projdim k,\forall M\in{_V\calM}\]
	but then
	\[\projdim(M)=\projdim(k\otimes M)\le \projdim(k)\]
	by lemma~\ref{lem-tensorproj}
\end{prf}
\begin{ex}
	Today's example: Quantum group $V=\mathcal{O}_q(GL_2)$. Fix some $q\in k^*$. Then as an algebra, 
	\[V=k\langle x_{11},x_{12},x_{21},x_{22},det_q^{-1}\rangle/(\text{relations})\]
	where the relations are given by 
	\begin{align*}
		x_{ij}x_{kl}=(\delta_{ik}+\delta_{jl})x_{kl}x_{ij}\quad\text{except when } i=k \text{ and } j=l\\
		x_{12}x_{21}=x_{21}x_{12}\\
		x_{22}x_{11}-x_{11}x_{22}=(q-q^{-1})X_{12}x_{21}\\
		det_q^{-1}det_q=det_q det_q^{-1}=1
	\end{align*}
	where $det_q=x_{22}x_{11}-q^{-1}x_{12}x_{21}x_{11}x_{22}-qx_{12}x_{21}$. One can show $det_q$ is central in $V$.

	Now for the coalgebra: $\delta(X)=X\otimes X$, so $\Delta(x_{ij})=\sum_{s=1}^2 x_{is}\otimes x_{sj}$, $\varepsilon(X)=I_2$ and
	\[S(X)=X^{-1}=\begin{pmatrix}
		x_{22}det_q^{-1} & -qx_{12}det_q^{-1}\\
		-q^{-1}x_{21}det_q^{-1} & x_{11}det_q^{-1}
	\end{pmatrix}\]
\end{ex}
\begin{prob}
	Show that $\gldim(\begin{smallmatrix}\Z & \Q\\0 & \Q\end{smallmatrix})=2$ while $\gldim(\begin{smallmatrix}\Z & \Q\\0 & \Q\end{smallmatrix})^{op}=1$
\end{prob}

\subsection{Returning to Homological Algebra}
Let $A$ be an algebra, $M\in {_A\mathcal{M}}$. Recall the definitions of projective and global dimension.
\begin{rmk}
	The following are equivalent:
	\begin{enumerate}
		\item $\gldim A=0$
		\item $\gldim A^{op}=0$
		\item $A$ is semisimple Artininan
		\item $A=\oplus_1^b M_{n_i}(D_i)$ by Artin-Wedderburn where $D_i$ is a division ring.
	\end{enumerate}
\end{rmk}
\begin{rmk}
	\begin{enumerate}
		\item A PID has global dimension 1. 
		\item The free algebra $k\langle x_1,\dots,x_n\rangle$ has global dimension 1.
		\item The path algebra of a finite quiver has global dimension 1.
		\item $\mathcal{O}_q(GL_2)$ has global dimension 4.
	\end{enumerate}
\end{rmk}

From now on, let $V$ always stand for some Hopf algebra.
\begin{lem}
	${_V\mathcal{M}}$ is a tensor category. The action on $M\otimes N$ is the following:
	\[v\cdot (m\otimes n)=\sum (v_{(1)}\cdot m)\otimes (v_{(2)}\cdot n)\]
\end{lem}
\begin{lem}
	If $M\in {_V\mathcal{M}}$, then $V\otimes M\in {_V^V\mathcal{M}}$. As a corollary,
	\[V\otimes M\cong V\otimes (V\otimes M)^{cov}\]
	and as a consequence $V\otimes M$ is a free $V$-module.
\end{lem}

\begin{thm}
	Let $P\in {_V\calM}$ be projective. Then $P\otimes N\in {_V\calM}$ is projective for all $M\in{_V\calM}$
\end{thm}
\begin{prf}
	To see this, notice that $P$ is a summand of $V^{\oplus b}$. But then if $P\oplus Q\cong V^{b}$, then 
	\[[P\otimes N]\oplus[Q\otimes N]=[P\oplus Q]\otimes N=(V^b)\otimes N=(V\otimes N)^b\]
	and by the last lemma, this is free.
\end{prf}

\begin{lem}\label{lem-tensorproj}
	\[\projdim(M\otimes N)\le \min\{\projdim M,\projdim N\}\]
\end{lem}
\begin{prf}
	We only show that $\projdim M\otimes N\le \projdim M$. If $\projdim M=\infty$, we are done.
	If it is finite, take any minimal projective resolution of $M$ and tensor with $N$. This is a projective 
	resolution of $M\otimes N$ of length at most $\projdim M$. 
\end{prf}
\begin{rmk}
	The fact it is still exact holds because we can appeal to the vector space structure.
\end{rmk}

\brk

Some more results (mostly for finite dimensional algebras):
\begin{thm}[Radford `75]
	If $V$ is finite dimensional, then the antipode $S$ has finite order:
	\[S^d=\id_V\]
\end{thm}
\begin{thm}[Larson-Radford `88]
	Suppose $\ch k\ne 0$ (can be relaxed but gets more ugly). Then the following are equivalent:
	\begin{enumerate}
		\item $\gldim V=0$ ($V$ is semisimple Artinian)
		\item $\gldim V^*=0$
		\item $S^2=\id_V$
	\end{enumerate}
\end{thm}
\begin{thm}[Nichols-Zoeller `89]
	Let $V$ be finite dimensional. If $Q$ is a Hopf subalgebra of $V$, then $_WV$ and $V_W$ are free.
\end{thm}

\section{February 6, 2019}
First some history: in 1899, Maschke proves the following:
\begin{thm}
	Let $G$ be a finite group. Then $kG$ is semisimple (Artinian) if and only if $\ch k\nmid |G|$.
\end{thm}
Today, we will see an analog of Maschke for Hopf algebras:
\begin{thm}[``Maschke's Theorem'' (Larson-Sweedler `69)]
	Let $V$ be a finite-dimensional Hopf algebra. Then the following are equivalent:
	\begin{enumerate}
		\item $\gldim V = 0$ ($V$ is semisimple Artinian)
		\item $\varepsilon(\int^l)\ne 0$
		\item $\varepsilon(\int^r)\ne 0$
	\end{enumerate}
\end{thm}

Now say $V=kG$. Then $\int=\sum_{g\in G} g$ is both a left and right integral of $V$. This uses that $\varepsilon(g)=1$.

But then if $\varepsilon(\int)=\sum_G \varepsilon(g)=\sum_G 1=|G|\ne 0\Leftrightarrow \ch k\nmid |G|$.

\subsection{Example of the Day}
\begin{ex}[Kac-Paljutkin Algebra]
	$k=\C$ or any field with $\ch k\ne 2$. Then $V=k\langle x,y,z\rangle$ modulo the relations
	\begin{align*}
		x^2=y^2=1\\
		xy=yx\\
		xz=yz\\
		xy=xz\\
		x^2=\frac{1}{2}(1+x+y-xy)
	\end{align*}
	Then $V$ has a $k$-basis $\{1,x,y,z,xy,xz,yz,xyz\}$, so as an algebra $V\cong k^4\oplus M_2(k)$.

	Then $\Delta(x)=x\otimes x$, $\Delta(y)=y\otimes y$, $\Delta(z)=(1\otimes 1+1\otimes x+y\otimes 1-y\otimes x)(z\otimes z)$.

	And $\varepsilon(x)=\varepsilon(y)=\varepsilon(z)=1$ and $S(x)=x$, $S(y)=y$ and $S(z)=z$. Recall that although $S$ looks like the identity map, 
	$S$ is an \textbf{antihomomophism}. So actually $S^2=\id_V$.

	Claim:
	\[\int=1+x+y+z+xy+xz+yz+xyz\]
	then $\varepsilon(\int)=8\ne 0$, so $V$ is semisimple.

	Some nice facts:
	\begin{enumerate}
		\item $V^*\cong V$
		\item $V$ is the unique eight dimensional noncommutative, noncocommutative semisimple Hopf algebra.
		\item Let $Q_8$ be the quaternion group. Then $kQ_8$ is eight dimensional. Therefore $V\cong kQ_8$ as algebras, but \textbf{not as coalgebras.}
		The reason behind this is that $kG$ is \textit{always} cocommutative.
	\end{enumerate}
\end{ex}

\subsection{HW of the day}
\begin{prob}
	If $V$ is eight dimensional and noncommutative, then $V\cong k^4\oplus M_2(k)$ as an algebra. Thus if $G$ is nonabelian group of order 8, then $\C G\cong \C^4\oplus M_2(\C)$.
\end{prob}

\subsection{Proof of ``Maschke's theorem''}
Now we actually do the proof:
\begin{prf}
	(a) $\Rightarrow$ (b): Say the global dimension of $V$ is zero. Then $V$ is semisimple by Artin-Wedderburn and we have a decompositon
	\[V\cong \oplus_1^s M_{n_i}(D_i)\]
	and every minimal nonzero ideal has the form $M_{n_i}(D_i)$.

	Consider the ideal $k\int^l=I$ which is an ideal since $v\int^l=\varepsilon(v)\int^l\in I$. But then consider that if $v'\int^l\in I$, 
	\[v'\int^l v=\varepsilon(v')\int^l v\in k\int^l\]
	so in fact it is a two-sided ideal. Thus $k\int^l\cong M_{n_i}(D_i)$, so 
	\begin{align*}
		(k\int^l)^2&=k\int^l
		\varepsilon(\int^l)\int^l=(\int^l)^2&=a\int^l
	\end{align*}
	for some nonzero $a$. But then $\varepsilon(\int^l)\ne 0$.

	(b) $\Rightarrow$ (a): Now let $e=\frac{1}{\varepsilon(\int^l)}\int^l$. Then $e^2=e$ and $ke=k\int^l$ is a left $V$-module.
	But then $V=Ve\oplus V(1-e)$ so $ke=k\int^l$ is projective, so $\projdim_v k=0$.

	The last part follows by Lorenz-Lorenz -- $\gldim V=\projdim_V k$.
\end{prf}

\subsection{$\Ext$ groups}
We define these in the usual way through projective resolutions:

Let $A$ be an algebra. Then $_A\calM$ is an abelian cateogry, so $\Hom_A(M,N)\in\Vectk$.
\begin{defn}
	A (covariant) functor $F:{_A\calM}\to{_B\calM}$ is called \textbf{left exact} if, for any short exact sequence
	\[0\to L\to M\to N\to 0\]
	we have the exact sequence
	\[0\to F(L)\to F(M)\to F(N).\]

	Similar for right exact. If both, it's just exact.
\end{defn}
\begin{lem}
	For any $W\in{_A\calM}$, the functor $\Hom(W,-)$ is left exact.
\end{lem}
We also have that $\Hom_A(P,-)$ is exact iff $P$ is projective, so this gives us a nice characterization of projective objects in Abelian categories.
Then the category $\Ch({_A\calM})$ is the category of chain complexes in $A$ modules with chain maps.
\begin{lem}
	$\Ch({_A\calM})$ is an Abelian category.
\end{lem}
\begin{lem}
	A functor $F:{_A\calM}\to{_B\calM}$ can be extended to a functor between the appropriate chain complex categories.
\end{lem}

\subsection{Looking Forward}
On Friday, we may see the following result:
\begin{thm}
	Under some mild hypothesis (including finite dimensionality),
	\[\projdim_V k=\gldim V\]
\end{thm}
\end{document}