### 2019-07-30 MSRI ML Notes/MSRI_ML_notes.tex
Day 2

\documentclass[12pt]{article}

\usepackage{setspace}

\usepackage{amsmath, amsfonts, amssymb, graphicx, color, fancyhdr, lipsum, scalerel, stackengine, mathrsfs, tikz-cd, mdframed, enumitem, framed, adjustbox, bm, upgreek, xcolor, hyperref}
\usepackage[framed,thmmarks]{ntheorem}

%Replacement for the old geometry package
\usepackage{fullpage}

%Input my definitions
\input{./mydefs.tex}

%Shade definitions
\theoremindent0cm
\theoremheaderfont{\normalfont\bfseries} 
\def\theoremframecommand{\colorbox[rgb]{0.9,1,.8}}
\newshadedtheorem{defn}[thm]{Definition}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%%%%%%%%%%%%%%%%%%%%%% Customize Below %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

%header stuff
\setlength{\headsep}{24pt}  % space between header and text
\pagestyle{fancy}     % set pagestyle for document
\lhead{``Mathematics of Machine Learning'' Notes} % put text in header (left side)
\rhead{Nico Courts} % put text in header (right side)
\cfoot{\itshape p. \thepage}
\setlength{\headheight}{15pt}
\allowdisplaybreaks

% Document-Specific Macros
\DeclareMathOperator{\err}{err}
\DeclareMathOperator{\argmin}{argmin}

\begin{document}
%make the title page
\title{MSRI's \textit{Mathematics of Machine Learning} Summer School\vspace{-1ex}}
\author{Nico Courts}
\date{July 29-Aug 9, 2019}
\maketitle

\renewcommand{\abstractname}{Introduction}
\begin{abstract}
	These notes were the ones I took while attending MSRI's ``Mathematics of Machine Learning'' summer school
	at the University of Washington during the two weeks above.
\end{abstract}

\section{Statistical Learning}
This series of lectures was given by Robert Shapire from Microsoft Research.

The topic of the talks here are somewhere at hte nexus of supervised and statistical learning. The big idea concerning what we want o do is 
to learn how to do better in the future based on experience from the past. The hope is that these methods can be fully automated. This is a proper subset of AI, 
but definitely a core area of how we do it.

We are looking for easy to use, flexible algorithms that give us useful results. Ideally they would be interpretable, but 
this is usually a secondary result.

He believes that ML can help us to understand how learning happens in non-machine entities (cool!). For instance, one can consider the question of 
nature vs nuture and to ask the question ``what is simplicity and why is it useful?''

\subsection{The Problem}
In this course, we will be focusing on a single kind of (simple) learning problem: learning from example. Given a set of examples, 
we are looking for an algorithm that can classify instances given objects.

For instance, character recognition. We have a set of handwritten characters and the letter it depicts. Then we want to feed this 
into a learning algorithm that gives us a prediction rule, that given a handwritten character, outputs the predicted character.

We were then given a couple of examples to try. The particulars weren't important, but the takeaway here is 
\begin{itemize}
	\item We needed enough data to say something meaningful.
	\item We looked for a rule that fit the observations. We want the rule to be consistent or to contain few mistakes.
	\item We were looking for rules that were ``simple.'' Importantly, our notion of simplicity changed when we converted one set of numbers to binary.
\end{itemize}
The first condition is always the case: ``more is more.'' But the latter two fit into a tradeoff between fit and simplicity.

The main questions we will be trying to answer:
\begin{itemize}
	\item How much data do we need?
	\item How do we define simplicity?
	\item How much complexity do we need to represent our problem?
	\item What is the tradeoff mentioned above?
\end{itemize}

\subsection{Studying the Problem Mathematically}
We have to come up with a formal model for the problem so that we can do mathematics with it. The learning model should answer some basic questions:
\textit{What is the goal of learning?} and \textit{How is the learning happening?}.

We begin the description of the model:
\begin{defn}
	An \textbf{instance} (or sometimes informally \textbf{example}) $x$ is an object in a space called the \textbf{instance space} or \textbf{domain} $X$.
\end{defn}
\begin{rmk}
	To each instance we assign a \textbf{label} or \textbf{class}. In this course we will be simplifying to only consider two classes, denoted $0$ and $1$ or $+$ and $-$.
\end{rmk}
\begin{rmk}
	We work under an assumption that there is an underlying function 
	\[c:X\to\{0,1\}\]
	called the (target) \textbf{concept} that we are trying to learn. Later we will relax this assumption.
\end{rmk}
\begin{defn}
	We define the \textbf{hypothesis} to be (similar to the concept) a map 
	\[h:X\to\{0,1\}\]
	that holds our current beliefs.
\end{defn}
We write $\calC$ to denote the \textbf{concept class}, the class of all possible functions we are exploring (we assume we know a class to which $c$ belongs). We will further 
assume that our data were generated \textit{independently} and at random. It is limiting to make an assumption about the distribution from which these were drawn. So we will just say there is one.

\begin{defn}
	We define 
	\[\operatorname{err}_D(h)=\Pr_{x\sim D}[h(x)\ne c(x)]\]
	to be the (generalization) \textbf{error}.
\end{defn}

Then we obviously ask that $\operatorname{err}_D(h)<\varepsilon$ for some small $\varepsilon$, but furthermore we want that 
this happens with high probability:
\[\Pr[\operatorname{err}_D(h)\le\varepsilon]\ge 1-\delta\]
and in this case, we say that that the rule is probably approximately correct  (PAC).
\begin{defn}
	The class $\calC$ is \textbf{PAC-learnable} by $\calH$ (a hypothesis space) if there is an algorithm $A$ such that
	for all $c\in\calC$ and for all distributions $D$ and for all positive $\varepsilon$ and $\delta,$ then $A$ takes ``polynomial many''
	random variables $X_i\sim D$ which outputs $h\in\calH$ such that
	\[\Pr[\operatorname{err}_D(h)\le\varepsilon]\ge 1-\delta\]
\end{defn}
\begin{rmk}
	This is not an easy problem, per se, but it is one of the most simple classes of algorithms to work with.
\end{rmk}
\begin{ex}
	Let $X=\bbR$ and let $\calC$ be of the form $c_r$ for $r\in\bbR$ where 
	\[c(x)=\left\{\begin{array}{lr} 1, & x\ge r\\
	 0 & \text{otherwise}\end{array}\right.\]
	 so $\calC$ is the \textit{set of positive half lines}. Then if we have some data, we can set our hypothesis 
	 $h=c_b$ where $b$ is the leftmost positive value. Then if $c=c_r$, then $\operatorname{err}_D(h)$ has to do with the interval $[r,b]$. Let $b-r=\varepsilon$.

	 Now restrict attention to $X=[0,1]$ and notice that if we got a training point within $\varepsilon$ of $c$ (to the right), then we would have gotten a smaller difference. Thus
	 \begin{align*}
		 \Pr[\operatorname{err}_D(h)<\varepsilon] &\le \Pr[\text{no $x_i$ are in $[b,b+\varepsilon]$}]\\
		 &=\prod_1^m\Pr[x_i\in R]\\
		 &=(1-\varepsilon)^m\\
		 &\le e^{-\varepsilon m}
	 \end{align*}
	 using the fact that $1+x\le e^x$ for all $x$.
\end{ex}

\subsection{Sufficient conditions for learning}
Now we consider the case when our hypothesis space $\calH$ is finite.
\begin{thm}
	Let $A$ be an algorithm that finds a hypothesis $h_A\in\calH$ which is consistent with $m$ random training examples (as before)
	where
	\[m\ge \frac{1}{\varepsilon}(\ln|\calH|+\ln\frac{1}{\delta})\]
	then $\Pr[\operatorname{err}_D(h_A)>\varepsilon]\le \delta$.
\end{thm}
\begin{rmk}
	Note that we have completely done away with trying to get a handle on the concept class here.
\end{rmk}
\begin{rmk}
	Equivalently, with probability of $1-\delta$, 
	\[\operatorname{err}_D(h_A)\le\frac{\ln|\calH|+\ln|\frac{1}{\delta}|}{m}\]
\end{rmk}
The idea to take away here is that $\ln|\calH|$ is in some way measuring ``description length'' of a hypothesis: you have to describe how to distinguish the correct one from the wrong ones. Think popping these 
in a binary tree. Then the $m$ bound gives you some sort of way to manage the complexity of the algorithm.

\begin{thm}
	Assume we have $m\ge \frac{1}{\varepsilon}(\ln|\calH|+\ln|\frac{1}{\delta}|)$ examples. Then with probability $1\ge 1-\delta$, for all $h\in\calH$ if $h$ is consistent then $\operatorname{err}_D(h)\le \varepsilon$ (here we often write ``$h$ is $\varepsilon$-good.'')
\end{thm}
\begin{prf}
	We prove the converse: the probability of there being an $h\in\calH$ that is consistent and $\varepsilon$-bad is less than $\delta.$
	Notice that whether $h$ is consistent with the training set is a random variable (since it depends on choice of training set). Whether it is $\varepsilon$-bad is not random. So 
	if $\calB$ is the set of $\varepsilon$-bad hypotheses, then we really want 
	\[\Pr[\exists h\in\calB: h\text{ is consistent}]\le \sum_{g\in\calB}\Pr[h\text{ is consistent}]\]
using the union bound.

Now fix $g\in\calB$ and we want to compute the probability of it being consistent. That is, $h(x_i)=c(x_i)$ for all $i$. Since the samples are independent, this probability is 
\[\prod_1^m\Pr[h(x_i)=c(x_i)]\le (1-\varepsilon)^n\]

Therefore the original probability is 
\[|\calB|(1-\varepsilon)^m\le|\calH|e^{-\varepsilon m}\le \delta\]
\end{prf}

\subsection{When does this work when we have infinite hypothesis classes?}
We saw in the first example that infinite classes can still be PAC-learnable, but where is the line drawn?

To see this, consider classifications of finite subsets. But any choice of threshold between any two points on $\bbR$ will give the same classification! So really we only have 
something like 5 different hypothesis (in the form of a threshold function at least). Compare that with $2^m$ behaviors one could consider in general!

More formally, if we are given a set $S=\langle x_1,\dots,x_m\rangle$ of instances, then we can consider the collection of all $\langle h(x_1),\dots, h(x_m)\rangle$ for all $h\in\calH$. We call this set 
$\Pi_\calH(S)$ and define 
\[\Pi_\calH(m)=\max_{|S|=m}|\Pi_\calH(S)|\]
and call this the \textbf{growth function.} We can use this function to get some sort of handle on the complexity of the problem, 
and in fact it plays a similar role in an analogous theorem to the one we saw earlier:
\begin{prop}
	Give $m$ training examples, with probability $1-\delta$ for all $h\in\calH$ if $h$ is consistent then 
	\[\operatorname{err}_D(h)\le\calO\left(\frac{\ln\Pi_\calH(2m)+\ln\frac{1}{\delta}}{m}\right)\]
\end{prop}
\begin{rmk}
	The proof of this result can't quite continue in the same way as before because we need to select a training set \textit{before we can know the hypothesis space}, 
	so the dependence gets in the way. There is a fancier proof but we won't do it here (although there will be another proof later that is similar).
\end{rmk}
\begin{rmk}
	Notice that if $\Pi_\calH(m)$ is polynomial, this reduces nicely and as $m$ gets larger, the error drops. Thus learning is possible.
\end{rmk}

\subsubsection{More on the growth function}
There are at least some cases when $\Pi_\calH(m)\in\calO(m^d)$ for some constant $d$. In the worst case, $\Pi_\calH(m)=2^m$. In fact, these are the only two cases that can happen! Furthermore,
it ends up that these two cases correspond exactly to when learning is possible!

To see why, we need a new definition:
\begin{defn}
	A sample $S$ of size $m$ is \textbf{shattered} by $\calH$ if all possible behaviors/labelings are possible: that is, $\Pi_\calH(S)=2^m$.
\end{defn}
\begin{ex}
	Given $\calH$ to be the set of all closed intervals $[a,b]\subseteq\bbR$, any set of two points in $\bbR$ is shattered by $\calH$.
	But you cannot shatter a set of three points in $\bbR$.
\end{ex}
\begin{defn}
	The \textbf{Vapnik-Chervinenkis (or VC) dimension} is
	\[\operatorname{VCdim}(\calH)=\max\{|S|: S\text{ is shattered by }\calH\}\]
\end{defn}
\begin{rmk}
	So in the exercise above, $\operatorname{VCdim}(\text{intervals})=2$.
\end{rmk}
Another set of hypotheses are the linear threshold functions in $\bbR^n$, and the VC dimension here is $n+1$. If you fix your planes to be subspaces, the VC dimension is $n$.

\begin{rmk}
	A good heuristic measure for VC dimension is the number of parameters. There are pathological examples, but they are truly that.
\end{rmk}

We have a nice little lemma:
\begin{lem}[Sauer]
	If $d=\operatorname{VCdim}(\calH)$, then 
	\[\Pi_\calH(m)\le\sum_0^d\binom{m}{i}.\]
\end{lem}
\begin{rmk}
	A ``nicer'' upper bound leverages
	\[\sum\binom{m}{i}\le\left(\frac{em}{d}\right)^d\]
	whenever $m\ge d\ge 1$.
\end{rmk}

Plugging this into the bound on the error of a hypothesis, we get that (forgetting the relatively small log term) that the VC dimension $d$ gives us
a complexity measure---that is, it gives a bound on the error of a consistent hypothesis.

\subsection{Non-consistent Hypotheses}
The results so far have been great, but has always assumed that there is a consistent hypothesis to find. But what if our data is noisy? Then 
learning the concept is no longer about finding the concept that matches \textit{all} the given data.

Now we consider data $(x,y)\sim D$ to be a pair jointly sampled from some distribution $D$. We do basically the same 
thing to compute error:
\[\err_D(h)=\Pr_{(x,y)\sim D}[h(x)\ne y].\]
We are still trying to minimize the generalization error $\err(h)$ over $\calH$ given some sample $(x_i,y_i)$ sampled from $D$.
\begin{defn}
	Write the \textbf{empirical error} as 
	\[\widehat\err(h)=\frac{1}{m}\sum_1^m\delta_{h(x_i)\ne y_i}\]
\end{defn}
And then we can follow the route called \textit{empirical risk minimization} (ERM) that finds an empirical solution:
\[\hat h=\argmin_{h\in\calH}\widehat\err(h).\]

We want to get to the situation in which, with probability greater than $1-\delta$, for all $h\in\calH$,
\[|\err(h)=\widehat\err(h)|\le\varepsilon.\]
This result is a kind of \textbf{uniform convergence result} which is the kind of thing we are after. Obviously this is useful since it allows us 
to bridge the gap between imperfect information and the true concept.

Now $\delta_{h(x_i)\ne y_i}=1$ with probability $\err(h)$ and zero otherwise.

\begin{ex}
	A quick aside: Consider IID variables $Z_1,\dots,Z_m$ with $Z_i\in[0,1]$. Let $p=\bbE[Z_i]$ and then 
	let $\hat p=\frac{1}{m}\sum_i Z_i$ be the empirical mean.

	Then there is a result called \textbf{Hoeffding's inequality}, which claims
	\[\Pr[\hat p\ge p+\varepsilon]\le e^{-2e\varepsilon^2m}.\]

	Notice that what this says is that the empirical average converges to its expected value. So we can think of this 
	as a function converging
	\[f(Z_1,\dots,Z_m)\to\bbE[f(Z_1,\dots,Z_m)].\]
\end{ex}
This idea leads us to:
\begin{lem}[McDiarmid's Inequality]
	Suppose $f(z_1,\dots,z_m)$ is real valued such that chainging $z_i$ changes $f$ by at most $c_i$, and furthermore the $z_i$ are independent, but not necessarily identically distributed.
	Then you get a nice bound on the error (I didn't get it in time but I am sure it is online.)
\end{lem}

So if we let $Z_i=\delta_{h(x_i)\ne y_i}$ and let $p=\bbE Z_i=\err(h)$ and $\hat p=\widehat\err(h)$, then for any particular $h\in\calH$,
\[\Pr[|\err(h)-\widehat\err(h)|\ge \varepsilon]\le 2e^{-2\varepsilon^2 m}.\]
So when $\calH$ is finite, we can use the union bound to say 
\[\Pr[\exists h\in\calH: |\err(h)-\widehat\err(h)|\ge \varepsilon]\le 2|\calH|e^{-2\varepsilon^2 m}\]
then setting the RHS equal to $\delta$,  we get that with probability $1-\delta$, for all $h\in\calH$,
\[|\err(h)-\widehat\err(h)|\le\calO\left(\sqrt{\frac{\ln|\calH|+\ln\frac{1}{\delta}}{m}}\right)\]
or in another (slightly weaker but sometimes more suggestive) form:
\[\err(h)\le\widehat\err(h)+\calO\left(\sqrt{\frac{\ln|\calH|+\ln\frac{1}{\delta}}{m}}\right)\]
\begin{rmk}
	Notice that this really (finally) ties together how several aspects impact learning: it relates the true error to the sample size, the complexity class of the concept, 
	and the consistency of the hypothesis with the sample data.
\end{rmk}

\begin{rmk}
	Notice that as the complexity of your hypotheses increase, we expect the empirical error to decrease, but if we assume that 
	the above is truly a good bound for $\err(h)$, we would expect that the true error will begin by dropping, but eventually the growth term from $\ln|\calH|$
	will overtake things and cause the error to rise again.

	This is one way to think about the concept of \textbf{overfitting}: the complexity of your hypothesis functions can incorporate more and more pathological functions that are less likely to be realistic.
\end{rmk}

\subsection{Another Complexity Measure}
Suppose we have a labelled sample $S$ consisting of $(x_i,y_i)$, where we can now assume the label space is $\pm 1$.
Then if we have a hypothesis, we can compute the training error $\widehat\err(h)$:
\[\widehat\err(h)=\frac{1}{m}\sum_i\delta_{h(x_i)\ne y_i}=\frac{1}{m}\sum_i \frac{1-y_ih(x_i)}{2}=\frac{1}{2}-\frac{1}{2}\left(\frac{1}{m}\sum_i y_ih(x_i)\right)\]
and the term in parentheses is another quantity we can consider which is just a constant away from training error.

So we can consider computing 
\[\bbE_{\mathbf\sigma}[\max_{h\in\calH}\sum_i\sigma_ih(x_i)]\]
where $\sigma_i$ are $\pm 1$ equally likely. In the case that $|\calH|=1$, we get that this value is zero and when $\calH$ shatters $S$ (the set of all $m$-tuples of $\pm 1$), 
the value is 1. These serve as the two extremes, obviously.

We would like to work with this towards describing convergence (and rates thereof) towards the generalization error.
To do this, we need to broaden the context slightly. Let $f:Z\to\bbR$ (standing in for the hypothesis), a space $\calF$ ($\calH$), and define the \textbf{(empirical) Rademacher complexity} to be 
\[\hat R_\scrF(S)=\bbE_\sigma\left[\sup_{f\in\scrF}\frac{1}{m}\sum_i\sigma_if(z_i)\right]\]
and the expected Rademacher complexity is 
\[R_\scrF(m)=\bbE_S[\hat R_\scrF(S)]\]

Then we want to show that for all $f\in\scrF$ that $\hat\bbE_S[f]\to\bbE[f]$. Specifically,
\begin{thm}
	Let $\scrF$ be a family of functions $f:Z\to[0,1]$ and let $S$ be a collection of $z_i$ where $z_i\sim D$. Then with probability $1-\delta$,
	and for all $f\in\scrF$,
	\[\bbE[f]\le\hat\bbE_S[f]+2\cdot\Psi+\calO\left(\sqrt{\frac{\ln(1/\delta)}{m}}\right)\]
	where $\Psi$ is either $\hat R_\scrF(S)$ or $R_\scrF(m)$.
\end{thm}
\begin{prf}
	Let $\Phi(S)=\sup_{f\in\scrF}(\bbE[f]=\bbE_S[f])$. Then first we show (via McDiarmid) that 
	\[\Phi(S)\le \bbE_S[]\Phi(S)]+\calO\left(\sqrt{\frac{\ln 1/\delta}{m}}\right).\]
	So then we have reduced the problem with finding some bound for $\bbE_S[\Phi(S)]$.

	Then let $S'$ to be a hypothetical second sample (called the ``ghost sample''). Compute 
	\[\bbE_S[\Phi(S)]=\bbE_S[\sup_f(\bbE[f]-\hat\bbE_S[f])]\le \bbE_{S,S'}[\sup(\hat\bbE_{S'}[f]-\hat\bbE_S[f])]\]
	There is something to prove in the last line there but he didn't show us. He said there was a slide somewhere. :)

	For the next step, go through the two samples $S$ and $S'$ one index at a time and for each $i$ with equal probability either swap $z_i$ and $z_i'$ or do nothing.
	Call the resulting samples $T$ and $T'$. But then the following two expressions have the same distribution:
	\[\hat\bbE_{S'}[f]-\hat\bbE_S[f]\quad\text{and}\quad \hat\bbE_{T'}[f]-\hat\bbE_T[f]\]
	and so if we let $\sigma_i$ denote the random variable for whether we swapped at the $i^{th}$ step ($+1$ if we swapped $z_i$ and $z_i'$ and $-1$ otherwise),
	the latter is 
	\[\frac{1}{m}\sum_i\sigma_if(f(z_i')-f(z_i))\]
	so we get 
	\[\bbE_{S,S'}[\sup_f(\hat\bbE_{S'}-\hat\bbE_S[f])]=\bbE_{S,S'}\left[\sup_f\left(\frac{1}{m}\sum_i\sigma_if(f(z_i')-f(z_i))\right)\right]\]

	Finally we can show (another slide) that this value is at most $2R_\scrF(m)$ and then use McDiarmid's once more to show that the expected Rademacher complexity is close enough to that of the empirical RC.
\end{prf}

Now the whole point was to show that $\err(h)\approx \widehat\err(h)$. One can show the empirical Rademacher complexity of the class of loss functions comprised of indicator functions $l(x,y)=\delta_{h(x)\ne y}$ is twice that of $\calH$ itself.
You can easily plug this in to get $\bbE[l_h]$. Then you can get what you want.

Now if $|\calH|<\infty$, (careful, he made a notation error earlier so I am switching subscripts and arguments in Rademacher complexities--they are still the same)
\[\hat R_S(\calH)\le\sqrt{\frac{2\ln|\calH|}{m}}.\]

If the space is infinite, then we choose a representative hypothesis for each behavior seen in applying all hypotheses to $S$. Call this space $\calH'$ and notice that $|\calH'|=|\Pi_\calH(S)|.$
Then one can show that the empirical Rademacher complexity of $\calH'$ is the same as that of $\calH$!

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

\section{Convex Optimization}
This series was given by Sebastien Bubeck from Microsoft Research.

\subsection{Fundamentals}
Let $K\subseteq \bbR^n$ be a convex set: that is one such that the straight line segment between two points in $K$ is contained entirely in $K$.
\begin{defn}
	The map $f:K\to\bbR$ is called \textbf{a convex function} if 
	\[f((1-\gamma)x+\gamma y)\le (1-\gamma)f(x)+\gamma f(y).\]
\end{defn}
\begin{rmk}
	Then if $f$ is differentiable, this implies
	\[\frac{f(x+\gamma(y-x))-f(x)}{\gamma}\le f(y)-f(x)\]
	and as $\gamma\to 0$, the LHS tends to $\nabla f(x)\cdot (y-x)$. Thus
	\[f(y)\ge f(x)+\nabla f(x)\cdot(y-x).\]
\end{rmk}

Then the goal of this course is to ``find'' $\operatorname{argmin}_{x\in K}f(x)$. A question we ask ourselves: how are $f$ and $K$ specified? Mostly in these lectures we will be 
focusing on \textit{simple} $K$ and functions $f$ such that $\nabla f(x)$ can be computed (for any $x$).

\subsection{Examples in machine learning}
\subsubsection{Regression}
Here we have a dataset: $(a_i,y_i)\in\bbR^n\times\bbR$ are the elements. Then we want to find a rule that assigns $y$ to $\mathbf a$. We focus on \textit{linear} rules.
That is, we are looking for a linear space in $\bbR^{n+1}$ that approximates the ``true'' values.

Tentatively we let our function be $a\mapsto x\cdot a$, parameterized by $x\in\bbR^n$. Then for each $x$ we can evaluate how the tentative function
fits our data. More specifically:
\[\frac{1}{m}\sum_1^m l(x\cdot a_i,y_i)\]
where $l$ is some loss function.

There are several loss functions one might consider:

\noindent\textbf{Least Squares}: 
\[l(u,v)=(u-v)^2\] 
The upshot here, however, is that the evaluation function above in this case is convex!
The idea here is that we are applying a linear function to a convex function. The other nice thing is that if you make a modeling assumption about the 
distribution underlying the label distribution, nice things happen.

Specifically, assume that $y\sim\calN(a\cdot \underline x,\sigma^2)$ where $\underline x$ is the \textit{true value}.
Notice we haven't put any assumption on the draw distribution, just the relationship between the (fixed) $a_i$ and $y$ given the truth $x$. One can compute 
\[\frac{1}{2\pi\sigma^2}\exp(-\frac{1}{2\sigma}(y_i-a_i\cdot x)^2).\]

Thus the likelihood of the entire data set is the product:
\[\frac{1}{(2\pi\sigma^2)^{m/2}}\exp(-\frac{1}{2\sigma^2}\sum_1^m(y_i-a_i\cdot x)^2)\]
and so we recover the least squares loss function as the objective we want to minimize. That is, the maximum likelihood estimator minimizes least squares fit.

\subsubsection{Classification}
Now restrict $y\in\{\pm 1\}$. A natural loss is 
\[l(u,v)=\delta_{\operatorname{sign}(u)\ne\operatorname{sign}(v)}.\]
This is a problem for this class since it is non-convex.

So instead we define \textbf{support vector machines} using the loss 
\[l(u,v)=\max(0,1-uv)\]
which is convex

Another example is the \textbf{logistic loss}
\[l(u,v)=\log(1-e^{-uv})\]
which you can see is a smooth upper bound on the 0-1 loss we started with. Minimizing this function is equivalent to minimizing the MLE for the \textit{logistic model:}

Denote $p(a)=\bbP(+1|a)$. Then we assume that the ``scale'' of our probability is given by a linear function:
\[\log\left(\frac{p(a)}{1-p(a)}\right)=\underline x\cdot a\Leftrightarrow p(a)=\frac{1}{1+e^{-\underline{x}\cdot a}}\]
and then the \textbf{logistic loss} is the negative log-likelihood of the logistic model.

\subsubsection{Graphical Models}
Given $a_1,\dots,a_m\in\bbR^n$, we want to infer the correlation structure of the variables (how are they all related).

Let's start from a (Gaussian) model: we assume we are drawing IID samples from $\calN(\mu,\underline\Sigma)$. A fact: if $(\Sigma^{-1})_{ij}=0$, then $x_i$ and $x_j$ are independent, conditioned on the rest.
Thus the goal is to estimate $\Sigma^{-1}$. So if the truth is $\Sigma$, the density of $a_1,\dots,a_n$ is 
\[\frac{1}{\sqrt{(2\pi)^n\det(\Sigma)}}\exp(-\frac{1}{2}(a_i-\mu)^T\Sigma^{-1}(a_i-\mu))\]
and by taking the negative log likelihood:
\[c+\frac{m}{2}\log\det\Sigma+\frac{1}{2}\sum_1^m(a_i-\mu)^T\Sigma^{-1}(a_i-\mu)=C+\frac{m}{2}\left(-\log\det\Sigma^{-1}+\operatorname{tr}(\Sigma^{-1}S)\right)\]
where $c$ is some constant and $S$ is the sample covariance matrix. Then to estimate $\Sigma^{-1}$ via MLE, we want to solve 
\[\operatorname{argmin}_{X>0}-\log\det(X)+\operatorname{tr}(XS)\]
where this is convex in $X$ (this is not obvious but we will see it in the problem session).

If you know that the true matrix is sparse, you may add a penalty term $\lambda \|X\|_1$. Similarly in regression you may add either $\|x\|_1$ or $\|x\|_2^2$.
In either case, these add curvature to your objective function. This may speed up the optimization process.

\subsubsection{Unsupervised Learning}
Finally we want to talk about \textit{clustering}. Notice that we have been slowly changing our dataset. First we dropped labels and now we are going to represent our dataset as a graph $G=(V,E)$ of interactions.
We would like to infer some partitioning of the vertices.

As a modelling assumption, we use the stochastic block model. There is a hidden partitioning $\sigma\in \{-1,1\}^|V|.$ $G$ is generated from the distribution
\[(i,j)\in E\text{ with probability }\left\{\begin{array}{lr}p & \sigma_i=\sigma j\\ q, \text{otherwise}\end{array}\right. \]
where we assume $q<p$. Furthermore we assume all connections are independent.

The the likelihood of $x\in\{-1,1\}^n$ is given as follows: let $A=(a_{ij})$ be the adjacency matrix of $G$. Then the likelihood is 
\[\prod_{ij}\left[\frac{1+x_ix_j}{2}\left(pa_{ij}+(1-p)(1-a_{ij})+\frac{1-x_ix_j}{2}(qa_{ij}(1-q)(1-a_{ij}))\right)\right]\]
this is non-convex! But we can look at a convex upper bound and we're good! Next time we'll talk about gradient descent.

\subsection{Gradient Descent}
This is the most simple algorithm we will study (also one of the most used ones). It is an iterative method where we define 
\[x_{t+1}=x_t-\eta\nabla f(x_t)\]
which dates back to the mid 1800's (likely by Cauchy). Notice that this is similar to Newton's method, but the latter requires inverting an $n\times n$ matrix 
and this was an attempt to avoid that difficult computation.

Why does this work? By Taylor's theorem, $f(y)\approx f(x)+\nabla f(x)\cdot (y-x)$. The idea here is that the negative $\nabla f(x)$ direction is in the direction of the minimum, so we just follow that.

Now if we have a convex function, where $f(x)-f(y)\le\nabla f(x)\cdot (x-y)$, we can set $y=x^\ast$ and define 
\[\delta(x):=f(x)=f(x^\ast)\le\nabla f(x)\cdot (x^\ast-x)=-\nabla f(x)\cdot(x^\ast-x)\]
then the idea is that progress in the $x^\ast-x$ direction (that is towards the target) is bounded below by $\delta(x)$.
That is to say: \textit{the rate of decrease of the instance to OPT is lower bounded by the suboptimal gap.}

There is a small problem, namely that we also take a (hopefully small) step away from the direct-line path from $x$ to $x^\ast$. To fix this, we
use continuous time analysis:
\[\frac{d}{dt}x(t)=-\nabla f(x(t))\]
which we call the \textbf{gradient flow.} Then 
\[\frac{d}{dt}\frac{1}{2}\|x(t)-x^\ast\|^2=\left(\frac{d}{dt}(x(t)-x^\ast)\right)\cdot(x(t)=x^\ast)=-\nabla f(x(t))\cdot (x(t)-x^\ast)\le -\delta(x(t))\]
and integrating over time:
\[\int_0^\tau \frac{1}{2}\|x(t)-x^\ast\|^2 \mathrm{d}t\le\int_0^\tau \delta(x(t))\mathrm{d}t\]
and using the FTOC and dropping the $\tau$ term (and normalizing):
\[\frac{1}{\tau}\int_0^\tau\delta(x(t))\mathrm dt\le \frac{1}{2\tau}\|x(0)-x^\ast\|^2:=\frac{r_0^2}{2\tau}\]
and so finally
\[\delta\left(\frac{1}{\tau}\int_0^\tau x(t) \mathrm{d}t\right)\le\frac{r_0^2}{2\tau}.\]

Of course this isn't a discrete process, so it's useless for coding! So we want to run gradient descent with step size $\eta$ for $T$ steps and approximate this integral with a finite sum from $1$ to $T$.
Then if we further impose that $f$ is Lipshitz ($\|\nabla f(x)\|\le L$) then the error in this approximation is at worst $\eta L^2$.
So if we are trying to measure the total error of our discretized algorithm from the true value, we get an error bound 
\[\frac{r_0^2}{T\eta}+\eta L^2\le \frac{r_0L}{\sqrt{T}}\]
for a ``good'' choice of $\eta=\frac{r_0}{L\sqrt{T}}$.
\subsubsection{Discrete Time Analysis}
\begin{thm}
	Let $g_1,\dots,g_T$ be arbitrary vectors in $\bbR^n$ and that $x_{t+1}=x_t-\eta g_t$. Then for any $x\in\bbR^n$,
	\[\sum_1^Tg_t(x_t-x)\\le \frac{\|x_1-x\|^2}{2\eta}+\eta\sum_1^T\|g_T\|^2\]
\end{thm}
\begin{cor}
	If $g_t=\nabla f(x_t)$ ($f$ convex) and $\|nabla f(x_t)\|\le L$, then
	\[f\left(\frac{1}{T}\sum_1^Tx_t\right)-f(x^\ast)\le\frac{1}{T}\sum_1^T(f(x_t)-f(x^\ast))\le\frac{1}{T}\sum \nabla f(x_t)(x_t-x^\ast)\le\frac{r_n^2}{2\eta T}+\eta L^2\] 
\end{cor}
The proof here is really just one line of computation but I missed it. Interestingly, the proof gives us \textit{equality} rather than inequality. I suppose that it's nice to use it as if it were a bound.

\subsubsection{Projective Gradient Descent}
A generalization is optimization with some constraint. So if we have a bounded convex set $K$, we just follow each step in gradient descent with a projection onto the closest point in $K$ (should we leave it).
This is only better for us since 
\begin{lem}
	$\|P_K(x)-z\|\le\| x-z\}$ for any $z\in K$.
\end{lem}
\begin{rmk}
	Notice that we are assuming here that both $K$ and $f$ are convex. Without convexity we definitely don't have this.
\end{rmk}

\subsection{Stochastic Gradient Descent}
We begin with a corollary of the theorem we just proved for standard GD:
\begin{cor}
	Say that $g_t$ is a random variable such that $\bbE[g_t|x_t]=\nabla x_t$ and that $\bbE[\|g_t\|^2]<B^2$.
	Then 
	\[\bbE[f\left(\frac{1}{T}\sum_1^T x_t\right)-f(x^\ast)]\le\frac{r_1^2}{2\eta}+\frac{\eta}{2}TB^2\le\frac{r_1B}{\sqrt{T}}\]
	given an optimal choice of $\eta$.
\end{cor}

So then we can use this result to define SGD, Stochastic Gradient Descent. Let $f(x)=\bbE_{(a,y)}l(x,(a,y))$.
The idea here is that we don't have the actual gradient of the loss function, but we have access (under the statistical learning framework) to a continuous stream of IID data.
So we define the \textbf{stochastic gradient}:
\[g_t=\nabla_x l(x_t,(a,t,y_t))\]
that is, compute the $x$ gradient of $l(x,(a_t,y_t))$ and evaluate at $x_t$. Then basically this pos right into the above corollary.

\begin{rmk}
	here the notation varies from statistical learning, so think of $x$ as our hypothesis (the thing we're optimizing over) and $(a,y)$ denotes a sample from the dataset.
\end{rmk}

Note that here you only get one pass on the data (although here you can directly control the generalization error as opposed to bounding it with testing error in the other class).

Instead, we can do multi-pass SGD where at each step we sample uniformly data points from your dataset (possibly having much more steps than there are data points) but then your function converges to an unbiased estimator for the test error rather than the generalization error.

\end{document}